\subsection{Mazes}

\subsubsection{Interface}

The core types that different parts of the program interface with are mazes, paths, maze solvers and maze generators. A maze stores its start and end positions, the width and height of the maze, and a two-dimensional array of cells that describe the layout, where each cell can be a wall or empty. A path is an array of actions where each action can either be checking whether a cell at a position is a wall or to move one step in a cardinal direction. A maze solver is a function that takes a maze as input and returns a path that solves that maze. A maze generator takes a width and a height as input and returns a randomly generated maze as output.

\subsubsection{Path Verification}

The verify\_path function takes a path and a maze as parameters and returns whether that path successfully moves from the start to the end of the maze without ever moving into a wall cell.

\subsubsection{Solver Wrapper}

To aid the creation of maze solvers, a solver\_wrapper function has been created that takes a function as input and returns a maze solver. The wrapper defines variables to keep track of the current position in the maze and the path it has taken to get there. The input function is then passed five arguments: the end position of the maze, a function to get the current position in the maze, a function to check whether a position lies within the maze, a function to check whether a cell at a position is a wall and a function to move one step in a cardinal direction. These functions also update the current position and path accordingly.

\subsection{Maze-Solving Algorithms}

\subsubsection{A*}

A* \cite{a-star-alg} is implemented as a maze solver using the solver wrapper. It will always find the shortest path and always halt.

\subsubsection{Depth-First Search}

DFS \cite{pkd} is implemented as a maze solver using the solver wrapper. It will not always find the shortest path but will always halt.

\subsubsection{Dijkstra's Algorithm}

Dijkstra's algorithm \cite{dijkstras-alg} is implemented as a maze solver using the solver wrapper. It will always find the shortest path and always halt.

\subsubsection{Maze Routing algorithm}

The maze routing algorithm \cite{maze-routing-alg} is implemented as a maze solver using the solver wrapper. It will not always find the shortest path and may not halt when the maze is not solvable.

\subsubsection{aMazing Language Implementations}

In addition to TypeScript, each maze-solving algorithm used is also implemented as strings executable by the aMazing Language in maze mode.

\subsection{Maze Generation}
\subsubsection{Recursive Division}
Recursive division is a maze generation algorithm that works by recursively dividing the board into smaller sections, adding horizontal and vertical walls with passage openings, and continues dividing until the sections reach a minimum size \cite{recursive-division-alg}.

\subsection{Component Overview}
React is a library that lets you build user interfaces with components. In a React app, a component is a reusable UI element defined as a function that returns JSX markup \cite{react-component}. This component overview provides a detailed explanation of the components.

\subsubsection{App}
The \texttt{App} component is the entry point to the React app.

\subsubsection{Context}
React’s Context API is used to create contexts that allow components to share state and stateful logic without the need to explicitly pass props \cite{react-context}.

\paragraph{Context Providers} \

The \texttt{Editor} and \texttt{Board} components share information with the \texttt{Header} component. Context provides a way for the components to access and share data. A \texttt{GridProvider} component is defined to wrap the components that need access to the shared information. In \texttt{App}, the \texttt{GridProvider} wraps both the \texttt{Header} and \texttt{Board} components, allowing them to share the \texttt{grid} and \texttt{disabled} states, as well as the stateful logic, which includes the functions \texttt{setGrid} and \texttt{setDisabled} to manage them. State variables are created using a \texttt{useState} Hook \cite{react-hooks}.

\begin{verbatim}
    const [grid, setGrid] = useState<Grid>(makeGrid(0, 0));
    const [disabled, setDisabled] = useState<boolean>(false);
\end{verbatim}

Similarly, \texttt{EditorProvider} shares information about its state variables and stateful logic.

\paragraph{Custom Hooks} \

To access shared information, custom Hooks \texttt{useGrid} and \texttt{useEditor} are defined and can be imported into other components.

\begin{verbatim}
    import { useGrid } from "./GridContext.tsx";
    const { grid, setGrid, disabled, setDisabled } = useGrid();
\end{verbatim}

\subsubsection{Board}
The \texttt{Board} component is an interactive grid where the maze-solving algorithms are visualized.

\paragraph{Data Types} \

\texttt{Board} uses two important types:
\begin{itemize}
    \item \texttt{Node}: Represents a single cell in the grid with properties about its row and column indices, and optional Boolean properties for the start node, the end node, and walls (\texttt{isStart}, \texttt{isEnd}, \texttt{isWall}).
    \item \texttt{Grid}: Represents the entire grid as a 2D array of nodes, with references to the start and end nodes, and the number of rows and columns.
\end{itemize}

\paragraph{Integration with Context} \

The component uses \texttt{useGrid}, which provides the grid state and a function \texttt{setGrid} to manage the grid state and trigger a new render when the state changes.

The \texttt{disabled} state in the \texttt{Board} component controls whether user interactions are allowed. When the \texttt{disabled} state is set to true, the board receives the \texttt{disabled} class, which sets pointer events to none in CSS, preventing interactions with the board.

\begin{verbatim}
    <div
        ref={boardRef}
        className={clsx({"disabled": disabled}, "board")}
        onMouseUp={handleMouseUp}
    >
\end{verbatim}

\paragraph{Grid Initialization} \

When the \texttt{Board} is mounted, a \texttt{useEffect} Hook \cite{react-hooks} gets dimensions using \texttt{boardRef} and calculates the number of rows and columns based on the size of a 24-pixel cell. Refs, created with the \texttt{useRef} Hook \cite{react-hooks}, remember information without triggering new renders and gives access to DOM elements in React \cite{react-hooks}. The grid is created using \texttt{makeGrid}, and the initial grid is updated with the new grid. The effect will run only once on mount (when the component first renders) because \texttt{setGrid} is stable and does not change between renders. The grid is initially created in \texttt{GridProvider}, but because the \texttt{Board} component is not rendered at this point and the dimensions are based on the layout defined by CSS, the dimensions cannot be retrieved until after the component has been mounted. This is why \texttt{useEffect} is used here.

\paragraph{Interacting with the Grid} \

The \texttt{Board} component enables user interactions through mouse event handlers.

\begin{itemize}
    \item \texttt{handleMouseDown}: This function is called when a mouse button is pressed on a cell and checks if the clicked cell is a start node, an end node, or a regular node. Then, it assigns an action to \texttt{nodeRef} or \texttt{wallRef}.
    \item \texttt{handleMouseEnter}: This function is called when the mouse moves over a cell while the mouse button is pressed down. It uses helper functions to move the start node or end node and modifies walls.
    \item \texttt{handleMouseUp}: When a mouse button is released, it resets the references \texttt{nodeRef} and \texttt{wallRef} to null and updates the grid state to sync changes made during interactions.
\end{itemize}

\paragraph{Rendering} \

The grid is rendered as an HTML table, where each cell is a node in the grid and has a unique ID that can be retrieved with \texttt{getNodeID}. Using IDs helps when manipulating cells in the DOM.

The update logic decouples visual feedback from the underlying grid state for performance reasons. When the user interacts with the board, the component uses helper functions to manipulate the DOM by adding or removing CSS classes on the cells to reflect the changes visually. When an interaction is completed, \texttt{updateGrid} is called to update the grid state.

\paragraph{Helper Functions} \

\begin{itemize}
    \item \texttt{editWall}: Modifies the class attributes of a specific cell. If the cell is not a start or end node, it changes the \texttt{isWall} property of the node based on whether the current action in \texttt{wallRef} is to add or remove a wall. This function adds or removes the corresponding CSS class to visually update the cell.
    \item \texttt{moveNode}: Repositions of the start and end nodes. First, it ensures that the target cell is not a wall and is not occupied by the opposite node. Then, it moves the node by removing the relevant CSS class from the old cell and adding it to the new one, while also updating the \texttt{isStart} or \texttt{isEnd} properties of the nodes.
    \item \texttt{styles}: Uses the clsx library to conditionally return a string of CSS classes for a given node. This helps in applying different styles for the cells (e.g. walls, start, and end nodes) based on the properties of the node.
\end{itemize}

\paragraph{Visualizing Algorithms} \

Algorithms are visualized on the board with an asynchronous function \texttt{visualize}. Because the \texttt{Grid} type is not directly compatible with the \texttt{Maze} type, which is used by the maze-solving algorithms, a helper function \texttt{gridToMaze} is used to convert the \texttt{Grid} to a \texttt{Maze}. The reason for the asynchronous approach is to see the progression of the algorithm searching and drawing the path, as the DOM updates would otherwise happen instantly. Each visualization step is delayed by 10 milliseconds using \texttt{setTimeout}. The Promise API provides a \texttt{resolve} function, which is used to indicate that an asynchronous operation (the delay) is complete and allows the next step.

\subsubsection{Editor}
The \texttt{Editor} component is a simple code editor for users to implement their own algorithms using aMazing Language.

It consists of two textareas: one for users to write code and another for displaying error messages that would otherwise appear in the console. A helper function, \texttt{handleKeyDown}, prevents the default tab behavior to let users make tab indentations within the editor. It detects tab key presses using the \texttt{KeyboardEvent} type from React and updates the position of the cursor.

\subsubsection{Header}
The \texttt{Header} component interfaces with the \texttt{Board} and \texttt{Editor} components. It inherits states and stateful logic from \texttt{GridProvider} and \texttt{EditorProvider}. The \texttt{disabled} state is used in \texttt{Header} to disable buttons and apply different styles. The \texttt{code} state and \texttt{setCode} and \texttt{setLog} functions are used to manage the code editor. When running algorithms from the code editor, \texttt{code} is passed to \texttt{evaluate\_solver}.

The different maze-solving algorithms are displayed in the dropdown using an array of objects. Additionally, a Custom option allows users to write their own algorithm. A placeholder function is used in place of the maze solver, as it is not defined until the code is evaluated.

\begin{verbatim}
    { id: 4, name: "Custom", fn: (): never[] => [], impl: "" }
\end{verbatim}

Implementations of the available algorithms can be displayed in the dropdown when selected. This is handled by a \texttt{useEffect} Hook that updates the \texttt{code} state using \texttt{setCode}.

The code editor will capture errors that would otherwise display in the console. These errors are displayed within the editor itself by updating the \texttt{log} state using \texttt{setLog}.

\paragraph{Helper Functions} \

\begin{itemize}
    \item \texttt{clearSearch}: Clears previous searches and paths if there are any.
    \item \texttt{clearBoard}: Resets the entire board and updates the grid state with \texttt{setGrid}, if there are walls, visited nodes, or paths.
    \item \texttt{generateMaze}: Resets the board and generates a maze using recursive division and updates the grid state.
    \item \texttt{run}: Starts the visualization of the selected algorithm. If the code editor is open, it runs the code in the editor.
\end{itemize}

\subsubsection{Styling}
All components are styled with Tailwind CSS, which is a CSS framework that defines a list of utility CSS classes that can be used to build designs directly in the markup. Utility classes are included in the class attribute to define the style.

\begin{verbatim}
    <div className="flex items-center space-x-2">
        <div className="node start" />
        <span>Start Node</span>
    </div>
\end{verbatim}

Tailwind scans the components for utility classes and generates the necessary CSS \cite{tailwind-utility}. This project also includes a CSS file in which Tailwind is imported to allow the use of directives like \texttt{@apply} and \texttt{@theme}, which are used to customize and extend the framework \cite{tailwind-directives}. In the example below, the \texttt{@theme} directive is used to define a custom animation for the nodes.

\begin{verbatim}
    @theme {
        --animate-node: node 0.3s ease-in-out;
    }
\end{verbatim}

\subsection{aMazing Language Interpreter}

\subsubsection{Tokenizer}

The first step of interpreting a string is to convert it to an array of tokens. A token consists of its type, the text it represents, the line number it started on and the column number it started on.

The input string is read one character at a time, and that character is added to a string containing the current token. When whitespace is encountered or the current token string would become invalid, the current token string is matched against keyword, symbols, integer literal rules, and variable naming rules. If it matches any of them, a new token of that type is pushed to the resulting array, and the current token string is reset to begin accumulating the next token.

\subsubsection{Parser}

Second is to convert the token of tokens into an intermediate representation. The intermediate representation is a tree-like structure where each node contains the type of expression or statement it represents, as well as what other values and nodes it consists of.

The input array is treated as a stack from which tokens are consumed as subexpressions are parsed. The main parse function will look at the next token and depending on what it is, it will call different helper functions capable of parsing that specific type of statement. Those functions will in turn call different helpers to parse the expressions they consist of, which will call helpers to parse their subexpressions. Once all subexpressions are parsed, they can be combined into the full expression node and returned to be combined into a full statement node, and so on.

\subsubsection{Evaluator}

The evaluator will use the tokenizer and parser to get the intermediate representation of the input. It uses an environment frame to store a table for looking up the values associated with variable names, as well as which frame to fall back on if a name could not be found.

The predefined values are defined in a new environment frame before starting the evaluation of the intermediate representation in a new frame with that as fallback. Similarly to the parser, when evaluating statements, helper functions are called to parse the expressions that make them up, which in turn call helper functions that evaluate their subexpressions, and so on. Eventually, a function is returned that will evaluate the main function of the input using this same method.

\subsubsection{Maze-Solving Evaluator}

To evaluate in maze mode, the solver wrapper is used. The regular evaluator is used, but the functions passed by the solver wrapper are exposed to the program as additional predefined values. Additionally, the main function is called with the end position's coordinates as arguments. A tuple of a maze solver and an array of strings is returned. The array of strings represents the program's standard output and will be filled with the strings printed during execution after the maze solver is called.
